---
title: "Decision Trees and Random Forests"
author: "Marcin Mazurek"
date:  '2017-11-08'
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
---





#Classification trees 
##Przykładowy zbiór danych

###Metadata
http://archive.ics.uci.edu/ml/datasets/statlog+(heart)

#### Attribute Information: 

-- 1. age 

-- 2. sex 

-- 3. chest pain type (4 values) 

-- 4. resting blood pressure 

-- 5. serum cholestoral in mg/dl 

-- 6. fasting blood sugar > 120 mg/dl 

-- 7. resting electrocardiographic results (values 0,1,2) 

-- 8. maximum heart rate achieved 

-- 9. exercise induced angina 

-- 10. oldpeak = ST depression induced by exercise relative to rest 

-- 11. the slope of the peak exercise ST segment 

-- 12. number of major vessels (0-3) colored by flourosopy 

-- 13. thal: 3 = normal; 6 = fixed defect; 7 = reversable defect 

#### Attributes types 


Real: 1,4,5,8,10,12 

Ordered:11, 

Binary: 2,6,9 

Nominal:7,3,13 


#### Variable to be predicted 
 
Absence (1) or presence (2) of heart disease 

### Import data 
```{r}

library(ggplot2)


heart<-read.csv("heart.dat", sep=' ')

colnames(heart)<-c('age',
  'sex' ,
  'chest_pain_type' ,
  'blood_pressure' ,
  'cholestor',
  'sugar',
  'resting_electrocardiographic_results' ,
  'heart_rate', 
  'angina' ,
  'oldpeak', 
  'slope' ,
  'vessles' ,
  'thal',
  'disease')


head(heart)

#binary columns
bin_cols <- colnames(heart)[c(2,6,9)]
nom_cols <- colnames(heart)[c(7,3,13)] 

#
heart[bin_cols]<-lapply(heart[bin_cols], factor)
heart[nom_cols]<-lapply(heart[nom_cols], factor)

#target_columns
target_col<-'disease'
heart[target_col]<-lapply(heart[target_col], factor)
#losowe pomieszanie kolejnosci rekordow 

heart.shuffled<-heart[sample(nrow(heart)),]


#indeksy rozgraniczajce partycje (rekordy sa juz pomieszane)
idx1<-round(0.6*nrow(heart.shuffled))
idx2<-round(0.8*nrow(heart.shuffled))


train<-heart.shuffled[1:idx1,]
validation<-heart.shuffled[(idx1+1):idx2,]
test<-heart.shuffled[(idx2+1):nrow(heart.shuffled),]

predictors_string<-paste( c(bin_cols ,  nom_cols), collapse='+')
formula<-as.formula(paste('disease~', predictors_string))

###prior probabilities 

t<-table(heart$disease)
prop.table(t)
```

##Budowa drzewa klasyfikującego
###Biblioteka tree 

####Budowa drzewa 
```{r}
library(tree)

heart.tree <- tree(formula, train)
summary(heart.tree)
heart.tree

plot(heart.tree)
text(heart.tree, pretty=0)


```


####Błąd klasyfikacji 

```{r}
p_disease<-predict(heart.tree, test )
#Stosunek poprawnie zaklsyfikowanych do wszystkich 
accuracy<-mean(max.col(p_disease)==test$disease)
misclasification_rate<-1-accuracy

accuracy

```

####Przycinanie (prunning)
```{r}

#przycinanie drzewa: wykreślenie błędu w zależności od liczby liśći 
#FUN - kryterium przycianania, 
cv.heart.tree<-cv.tree(heart.tree, FUN = prune.misclass)
cv.heart.tree

par(mfrow=c(1,2))
plot(cv.heart.tree$size, cv.heart.tree$dev, type="b")
plot(cv.heart.tree$k, cv.heart.tree$dev, type="b")



pruned.heart.tree=prune.misclass(heart.tree, best=4)
summary(pruned.heart.tree)


plot(pruned.heart.tree)
text(pruned.heart.tree, pretty=0)


```
####Błąd klasyfikacji 

```{r}
p_disease<-predict(pruned.heart.tree, test )
#Stosunek poprawnie zaklsyfikowanych do wszystkich 
accuracy<-mean(max.col(p_disease)==test$disease)
misclasification_rate<-1-accuracy

accuracy

```


####Biblioteka rpart 

```{r}

library(rpart)
library(rpart.plot)


heart.tree <- rpart(formula, train ,method="class")
summary(heart.tree)
rpart.plot(heart.tree, uniform=TRUE)

printcp(heart.tree)
opt <- which.min(heart.tree$cptable[,'xerror'])
cp<-heart.tree$cptable[opt,"CP"]

#pruned<-prune(heart.tree,cp)

plot(heart.tree)
text(heart.tree)

```



##Random forest
```{r} 
library (randomForest)


rforest<-randomForest(formula, train, ntree=50)

p_disease<-predict(rforest, test )
#Stosunek poprawnie zaklsyfikowanych do wszystkich 
accuracy<-mean(p_disease==test$disease)
misclasification_rate<-1-accuracy

accuracy

```

#Regression trees
